---
title: "Ellen Cheng's Coding Exercise- with Paige Duffin's Additions"
output:
  html_document:
    df_print: paged
---

### NOTE: I retained all the original coding from the author (Ellen Cheng) and interdispersed my new coding for the "Tidyverse" exercise. Blocks of code and text are headed by either "<font color="blue">EC:</font>" or "<font color="red">PD:</font>" when written by the original author or myself, respectively. I also add "_pd" to the end of all my variables I create. 

### <font color="blue">EC:</font> 
```{r}
library(dslabs) # load package with required data
```

### <font color="red">PD:</font>
Download and load `tidyverse` and `skimr` packages. 
```{r}
# install.packages("tidyverse") 
# install.packages("skimr")
library(tidyverse)
library(skimr)
```

### <font color="blue">EC:</font> 
```{r}
help(gapminder) # help on gapminder data
str(gapminder) # gapminder variable data types
summary(gapminder) # summary statistics of the data
class(gapminder) # gapminder is a dataframe
```

### <font color="red">PD:</font>
Inspect `gapminder` with `glimpse` from `dplyr`.
```{r}
?glimpse # view documentation on glimpse
glimpse(gapminder)
```

### <font color="red">PD:</font>
Inspect `gapminder` with `skim` from `skimr`.
```{r}
?skim # view documentation on skim
skim(gapminder)
```

### <font color="blue">EC:</font>  
```{r}
# Work with the Africa data ----
africadata <- subset(gapminder, continent == "Africa") # assign only the African countries to a new variable called africadata
str(africadata)
summary(africadata)
```

### <font color="red">PD:</font>
**Extract only the african countries using `filter` and `select` functions.**
Reminding myself of table structure:
```{r}
head(gapminder)
```
So the column I want is `continent` and the level I want is `Africa`

```{r}
africa_data_pd <- filter(gapminder, continent == "Africa")
head(africa_data_pd)
```

### <font color="blue">EC:</font> 
```{r}
# subset with infant mort
subset_inf_life <- africadata[, c("infant_mortality", "life_expectancy")] # infant mortality is per 1000, life expectancy is in years
str(subset_inf_life) 
summary(subset_inf_life) # there are 226 NA's for infant mortality
```

### <font color="blue">EC:</font>  
```{r}
# subset with pop
subset_pop_life <- africadata[, c("population", "life_expectancy")] # population is the country population size
str(subset_pop_life) # large range of country pop sizes
summary(subset_pop_life) # there are 51 NA's for population size
```

### <font color="red">PD:</font> 
Keep only variables `infant_mortality`, `life_expectancy`, `population` and `country`.
```{r}
pop_ctry_infmort_lifeexp_pd <- select(africa_data_pd, country, population, infant_mortality, life_expectancy)
head(pop_ctry_infmort_lifeexp_pd)
```

### <font color="blue">EC:</font> 
```{r}
# Plots ----
# life expectancy as a function of infant mortality
with(subset_inf_life, plot(life_expectancy ~ infant_mortality, main = "Life expectancy as a function of infant mortality", ylab = "Life expectancy (yrs)", xlab = "Infant deaths per 1000")) # negative correlation. Each streak is a different country.
```

### <font color="red">PD:</font>
**Plotting life expectancy as a function of infant mortality**
```{r}
library(ggplot2)
ggplot(pop_ctry_infmort_lifeexp_pd, aes(x=infant_mortality, y=life_expectancy, color=country)) +
  geom_point(size=2, shape=23, show.legend = FALSE) +
  ggtitle("Life expectancy by infant mortality in African countries") + 
  xlab("Infant mortality per 1000") +
  ylab("Life expectancy (yrs)")
```

### <font color="blue">EC:</font> 
```{r}
# life expectancy as a function of population size, using a log-scale on the x-axis
with(subset_pop_life, plot(life_expectancy ~ population, log = "x", main = "Life expectancy as a function of population size", sub = "(Note that x-axis is on log-scale)", ylab = "Life expectancy (yrs)", xlab = "Country population size")) # Each streak is a different country. Positive correlation within country.
```

### <font color="red">PD:</font>
This time around, it wasn't specified whether or not we should log transform population size, so I'm going to do both: 

**Plotting life expectancy as a function of population**
```{r}
ggplot(pop_ctry_infmort_lifeexp_pd, aes(x=population, y=life_expectancy, color=country)) +
  geom_point(size=2, shape=23, show.legend = FALSE) +
  ggtitle("Life expectancy by population size in African countries") + 
  xlab("Population size") +
  ylab("Life expectancy (yrs)")
```

### <font color="red">PD:</font>
**Plotting life expectancy as a function of log(population)**
```{r}
ggplot(pop_ctry_infmort_lifeexp_pd, aes(x=population, y =life_expectancy, color=country)) +
  geom_point(size=2, shape=23, show.legend = FALSE) +
  scale_x_continuous(trans='log') +
  ggtitle("Life expectancy by population size in African countries") + 
  xlab("log(population size)") +
  ylab("Life expectancy (yrs)")
```


### <font color="blue">EC:</font>  
```{r}
# More data processing ----
(missing_inf_mort <- unique(africadata$year[is.na(africadata$infant_mortality)])) # years with missing data for infant mortality
```

### <font color="blue">EC:</font>  
```{r}
# subset of Yr2000 data
subset_yr2000 <- subset(africadata, year == 2000)
str(subset_yr2000)
summary(subset_yr2000)
```

### <font color="red">PD:</font>
Make data subset that pulls out data for only African countries and only the year 2000.
```{r}
africa_data_yr2000_pd <- filter(africa_data_pd, year == "2000")
head(africa_data_yr2000_pd)
```

### <font color="blue">EC:</font> 
```{r}
# More plotting ----
# For Yr2000 ONLY, life expectancy as a function of infant mortality
with(subset_yr2000, plot(life_expectancy ~ infant_mortality, main = "Yr 2000: Life expectancy as a function of infant mortality", ylab = "Life expectancy (yrs)", xlab = "Infant deaths per 1000")) # negative correlation
```

### <font color="blue">EC:</font> 
```{r}
# For Yr2000 ONLY, life expectancy as a function of population size, using a log-scale on the x-axis
with(subset_yr2000, plot(life_expectancy ~ population, log = "x", main = "Yr 2000: Life expectancy as a function of population size", sub = "(Note that x-axis is on log-scale)", ylab = "Life expectancy (yrs)", xlab = "Country population size")) # No apparent correlation
```

### <font color="blue">EC:</font>  
```{r}
# Linear models (using Yr2000 data) ----
#  linear model predicting life expectancy as a function of infant mortality
summary(fit1 <- with(subset_yr2000, lm(life_expectancy ~ infant_mortality))) # p = 2.83E-8, so highly significant
```

### <font color="blue">EC:</font> 
```{r}
#  linear model predicting life expectancy as a function of population size
summary(fit2 <- with(subset_yr2000, lm(life_expectancy ~ population))) # p = 0.616, so not statistically significant at alpha = 0.05
```

### <font color="red">PD:</font>
Create a ggplot for life_expectancy as a function of infant_mortality in African countries only in the year 2000, with a linear model fit added. 
```{r}
ggplot(africa_data_yr2000_pd, aes(x=infant_mortality, y=life_expectancy, show.legend = FALSE)) +
  geom_point(size=2) +
  ggtitle("Life exp. by infant mortality in African countries in 2000") + 
  xlab("Infant mortality per 1000") +
  ylab("Life expectancy (yrs)") +
  geom_smooth(method='lm')
```


